Rewind: Automatically Generated\\Digital Memories from Mobile Geolocation Data

### 2016-09-20 paper/uist_2016/paper.tex
user study pretty much finished

\section{User Study}
\subsection{Research Questions}
% heather - is "digital memory" the right phrase for describing Rewind? Isn't rewind a medium for providing the cues to retrieve exisiting memories rather than creating new memories?
% talk about both within-subject and between-subject comparisons 
% within-subject -> memorability -> efficacy of semantic memory retrival vs episodic memory retrival -> video is better for both sementic and episodic
% between-subject -> memorability -> what type of users benefit the most from episodic memory retrival using Rewind; what kind of trip is Rewind best at capturing - (least frequently visited trip? most frequently visited trip? trip that's far away from home?)
The Rewind system enables us to ask questions about people's attitudes and behaviors surrounding digital memories that were not possible to answer before due to the scarcity of comprehensive data about where a person has been. To understand what makes a digital memory meaningful to a person, we performed a within-subjects comparison of Rewinds and static photos. By comparing Rewinds to today's most prevalent form of digital memory media, we can distill what features make a Rewind meaningful and representative as a memory of a trip.

Our research questions examine the effectiveness of \textit{Rewind} videos through three virtues of a memory: memorability, believability, and desirability. % heather - do we need a reference for the three virtues?

% heather - I changed this part...
\begin{enumerate}
\item Memorability: How would \textit{Rewind} facilitate the retrival of semantic and episodic memory? What types of users are most likely to recall their memories using \textit{Rewind}?
\item Believability:  How accurate are \textit{Rewind} artifacts in the eyes of the users' ? 
\item Desirability: What factors are inherent to users' decisions in preserving artifacts created by \textit{Rewind}?
\end{enumerate}

% heather - where to put this part or do we need a summary like this at all?
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Our user study explored the following five metrics that could potentially influence the aforesaid three measurements of \textit{Rewind} vidoes. These five metrics are:
\begin{enumerate}
\item GPS usage: how self-identified regular GPS users and non-regular GPS users differ in their responses to \textit{Rewind} videos.
\item Traveler type: how self-identified frequent and non-frequent travelers report to the three measurements of \textit{Rewind} videos differently.
\item Companions: how traveling companionship and solitude affect users' viewpoints on \textit{Rewind} videos.
\item Distance: how traveling distances influence users' perceptions of \textit{Rewind} videos.
\item Visit frequency: how visit frequency of a particular place influences users' reporting on \textit{Rewind} videos.
\end{enumerate}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Participants}
To recruit a diverse local population, participants for the user study were recruited online through local forums (Craigslist classifieds and local Reddit communities). Offline, ads were posted in cafes and other small businesses in the the city. Of the 21 total participants, 13 were identified as men and 8 were identified as women, with age ranged from 20 to 50 years old. Participants were mainly local residents, with occupations including librarian, medical researcher, office worker, IT consultant, student, musician, and comedian.
% Jeff notes, comment this out because it's not remarkable
% A fifth of the participants were undergraduate or graduate students.

\subsection{Procedure}
Participants underwent a two-week tracking period followed by a mixed-methods user study which investigated the memorability, believability, and desirability of \textit{Rewind} artifacts produced from the two weeks of location data.

@@

% heather - Rewind of the trip or Rewind video of the trip?
The participant was then shown a Rewind of the trip that they were on when the previous location was encountered. While the Rewind was playing, the participant could hover over the Rewind panel and scrub backwards and forwards horizontally through the images that composed the video. This allowed participants to ``walk through'' their memory at their own pace. The participant was asked the following questions while the Rewind played: Do you remember these locations? Do you remember this particular trip? Can you name some of the streets? Can you name what is around these locations, such as buildings, landmarks, or other places? Would you want to keep this video? Why or why not? Were you traveling alone? If you remember these locations, is the video accurate to your memory? Why or why not? What was different?

In addition to asking questions, we observed the users as they examined their generated Rewinds, since they could passively watch the Rewind or actively scrub through the individual images that made up the Rewind. The user study took between 30 minutes to an hour for each participant, and participants were compensated \$30 for their time at the end of the study.

% heather - ...that they did not carry the device on their person. - grammatical error?
Data from three participants was not included in the results of this study. One of the participants' data points consisted of only one location. The participant used a mobile device provided by the study, so it is likely that they did not carry the device on their person. Another participant did not complete the in-person interview portion of the user study due to health and scheduling reasons. The last participant only collected three days of data because they had their phone replaced during the tracking period, and did not reactivate Google's location reporting services on their new device. This participant completed the in-person interview portion of the user study, but we did not include the data gathered during this participant's session in the results.

% heather - need to be rephrased perhaps? Should we give a high level summary of all the relationships we've found? More analysis to be done for photo vs. videos, another table?
% semantic vs episodic memory
\section{Findings}
upper-bound estimates of event-related memories. 
The findings of our user study center on the participants’ relationship to the Rewinds generated from their location data. We focus on the participants’ judgments of their Rewinds’ memorability, believability, and desirability in comparison to their judgments of the photos paired with the Rewinds. We also discuss the participants' interactions with and reactions to their Rewinds. The results use data from the user studies conducted with 12 men and 6 women. Half of the participants identified themselves as frequent travelers. Additionally, half of the participants were regular GPS users, that is people who used GPS even in locations already familiar to them.

\subsection{Within-subject study}
\subsubsection{Memorability}
Within-subject comparison in our study reveals that participants are significantly better at recalling semantic memory of a particular location with \textit{Rewind} videos than with static street-view images, $\chi^2(1, 356) = 96.987$, $p < 0.01$. When being presented with a street-side image retrieved based on their location data, participants were able to recognize places visited 50\% of the time; yet recognition rate of locations reached 97\% when participants started playing accompanied \textit{Rewind} video, with a significant increase of 47\%. It is perhaps not too surprising as cinematic media like \textit{Rewind} video would provide users with more contextual meanings and retrival cues for semantic processing, as more than one participant have expressed their contentment after wathing \textit{Rewind} videos by claiming ``I recognize the photo now!'' when their initial attempts with static images were fruitless.

It could possibly be inferred that participants are likely to benefit more from \textit{Rewind} videos than static street-view images at recalling episodic memory. Although participants were not explicitly asked about trip rememberance when presented with static images, location rememberance from static images may serve as an upper-bound estimation of how much an individual could recall about a particular trip. - tbc
\subsubsection{Belivability and Desirability}
Although not statistically significant,

\subsection{Between-subject study}

\begin{table*}[t]
	\centering
	\caption{My caption}
	\small
	\label{within-subject-table}
	\begin{tabular}{c|c|c|c|c|c|c|c|c|}
		\cline{2-9}
		& \multicolumn{4}{c|}{Memorability} & \multicolumn{2}{c|}{Belivability} & \multicolumn{2}{c|}{Desirability} \\ \cline{2-9} 
		& \begin{tabular}[c]{@{}c@{}}video\\  location\end{tabular} & \multicolumn{2}{c|}{\begin{tabular}[c]{@{}c@{}}photo\\ location/trip\end{tabular}} & \begin{tabular}[c]{@{}c@{}}video\\ trip\end{tabular} & photo & video & photo & video \\ \hline
		\multicolumn{1}{|c|}{N} & 177 & \multicolumn{2}{c|}{179} & 177 & 86 & 128 & 179 & 177 \\ \hline
		\multicolumn{1}{|c|}{\%} & 96.61\% & \multicolumn{2}{c|}{49.72\%} & 53.67\% & 77.91\% & 79.69\% & 10.61\% & 14.69\% \\ \hline
		\multicolumn{1}{|c|}{p-val} & \multicolumn{2}{c|}{\begin{tabular}[c]{@{}c@{}}$\chi^2(1, 356) = 96.987$\\ p = 0.0001\end{tabular}} & \multicolumn{2}{c|}{\begin{tabular}[c]{@{}c@{}}$\chi^2(1, 356) = 0.41$\\ p = 0.2611\end{tabular}} & \multicolumn{2}{c|}{\begin{tabular}[c]{@{}c@{}}$\chi^2(1, 214) = 0.02$\\ p = 0.4434\end{tabular}} & \multicolumn{2}{c|}{\begin{tabular}[c]{@{}c@{}}$\chi^2(1, 356 ) = 0.995$\\ p = 0.1593\end{tabular}} \\ \hline
	\end{tabular}
\end{table*}
% 2nd table %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

% shall we change photo/rewind video rememberance to location/trip rememberance?
\subsection{Memorability}
Rewinds are meant to engage a person's memory. To find out more about what features make a digital memory memorable, we examined participant responses to whether or not they were able to remember the location shown in a photo or recall the trip portrayed by a Rewind. Photo memorability is determined by whether or not the participant remembered the location depicted in a photo. Because a Rewind contains many locations, it would be unfair to consider the memorability of Rewind based on whether participant remembers any locations in the Rewind. So, Rewind memorability is determined by whether or not the participant remembered and could distinguish the particular trip portrayed in a Rewind. In this section, we will look at differences in photo and Rewind memorability based on the variables of distance from the participant's home, visit frequency, presence of travel companions, regular GPS usage, and whether the participant considered themselves a frequent traveler.

% heather - didn't do the stats for video location vs photo locations, not sure if it is necessary? - photo vs video stats?
% will add this part back in within-group comparison

@@

@@

\subsubsection{Distance}
% only affect location rememberance
% 0 - 100%, 1 - 75%, 2 - 72%, 3 - 44%, 4 - 52%, 5 - 10%, $\chi^2(5,179) = 18.106$, p < 0.01
The distance away from the participant’s home had differing effects on photo memorability and Rewind memorability (Figure~\ref{fig:distancememory}). Distances ranged from 0 to 206,155 meters. For analysis, these distances were logarithmically scaled and normalized into six distance groups to mimic human categorization of distances into generalizations like ``close,'' ``far,'' or ``very far.'' When normalized: Group 0 corresponded to a range of 0--10 meters from the participant's home location; Group 1 corresponded to a range of 10--100 meters; Group 2 corresponded to a range of 100--1,000 meters; Group 3 corresponded to a range of 1,000--10,000 meters; Group 4 corresponded to a range of 10,000--100,000 meters and Group 5 corresponded to a range of 100,000--1,000,000 meters from the participant's home location.

>>>>

\section{User Study}
\subsection{Research Questions}
% heather - is "digital memory" the right phrase for describing Rewind? Isn't rewind a medium for providing the cues to retrieve exisiting memories rather than creating new memories?
% talk about both within-subject and between-subject comparisons 
% within-subject -> memorability -> efficacy of semantic memory retrival vs episodic memory retrival -> video is better for both sementic and episodic
% between-subject -> memorability -> what type of users benefit the most from episodic memory retrival using Rewind; what kind of trip is Rewind best at capturing - (least frequently visited trip? most frequently visited trip? trip that's far away from home?)
The Rewind system enables us to ask questions about people's attitudes and behaviors surrounding digital memories that were not possible to answer before due to the scarcity of comprehensive data about where a person has been. To understand what makes a digital memory meaningful to a person, we performed a within-subjects comparison of Rewinds and static photos. By comparing Rewinds to today's most prevalent form of digital memory media, we can distill what features make Rewinds, and digitally-mediated memories in general, meaningful and representative as a memory of a trip.

Our research questions examine the effectiveness of Rewinds compared to static photos through three virtues of a memory: memorability, believability, and desirability. % heather - do we need a reference for the three virtues? april - yeah that would be ideal!

% heather - I changed this part...
\begin{enumerate}% TODO TODO define semantic vs episodic memory
\item Memorability: How well do Rewinds facilitate the retrieval of semantic and episodic memory? What types of participants are most likely to recall their memories using Rewinds?
\item Believability:  How accurate are Rewinds in the eyes of the participants? 
\item Desirability: What factors are inherent to participants' decisions in preserving artifacts created by the Rewind system?
\end{enumerate}

\subsection{Participants}
To recruit a diverse local population, participants for the user study were recruited through online forums (Craigslist classifieds and local Reddit communities) and offline ads, which were posted in cafes and other small businesses in the the city. Of the 21 total participants, 13 identified themselves as men and 8 identified themselves as women, with ages ranging from 20 to 50 years old. Participants were mainly local residents, with occupations including librarian, medical researcher, office worker, IT consultant, student, musician, and comedian.
% Jeff notes, comment this out because it's not remarkable
% A fifth of the participants were undergraduate or graduate students.

\subsection{Procedure}
Participants underwent a two-week tracking period followed by a mixed-methods user study which investigated the memorability, believability, and desirability of Rewinds generated from the two weeks of location data.

@@

% heather - Rewind of the trip or Rewind video of the trip? april - we say 'rewind' when we mean rewind video, and if we want to refer to the system, we say 'the rewind system'
The participant was then shown a Rewind of the trip that they were on when the location in the photo was encountered. While the Rewind was playing, the participant could hover over the Rewind panel and scrub backwards and forwards horizontally through the images that composed the video. This allowed participants to ``walk through'' their memory at their own pace. The participant was asked the following questions while the Rewind played: Do you remember these locations? Do you remember this particular trip? Can you name some of the streets? Can you name what is around these locations, such as buildings, landmarks, or other places? Would you want to keep this video? Why or why not? Were you traveling alone? If you remember these locations, is the video accurate to your memory? Why or why not? What was different?

In addition to asking questions, we observed the participants as they examined their Rewinds. For instance, participants could passively watch the Rewind or actively scrub through the individual images that made up the Rewind. The user study took between 30 minutes to an hour for each participant, and participants were compensated \$30 for their time at the end of the study.

% heather - ...that they did not carry the device on their person. - grammatical error? april - nope (maybe out of fashion now though?), their person = their body
Data from three participants were not included in the results of this study. One of the participants' data points consisted of only one location. The participant used a mobile device provided by the study, so it is likely that they did not carry the device on their person throughout the day, but left it in one place. Another participant did not complete the in-person interview portion of the user study due to health and scheduling reasons. The last participant only collected three days of data because they had their phone replaced during the tracking period, and did not reactivate Google's location reporting services on their new device. This participant completed the in-person interview portion of the user study, but we did not include the data gathered during this participant's session in the results.

% heather - need to be rephrased perhaps? Should we give a high level summary of all the relationships we've found? More analysis to be done for photo vs. videos, another table?
% semantic vs episodic memory
\section{Findings}
% upper-bound estimates of event-related memories.  <-- april - found this without the beginning of its sentence. where does this fragment belong?
The findings of our user study center on the participants’ relationship to the Rewinds generated from their location data compared to the static photos. We focus on the participants’ judgments of their Rewinds’ memorability, believability, and desirability in comparison to their judgments of the photos paired with the Rewinds. We also discuss the participants' interactions with and reactions to their Rewinds. The results use data from the user studies conducted with 12 men and 6 women. Half of the participants identified themselves as frequent travelers. Additionally, half of the participants were regular GPS users, which we defined as people who use GPS navigation even in locations already familiar to them.

% heather - where to put this part or do we need a summary like this at all?
% april - moved to findings section from user study section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Our user study explored the following five metrics as potentially having an effect on memorability, believability, or desirability of digital memory media:
\begin{enumerate}
	\item Distance: how does the location's distance away from the participant's home influence the participant's perception of Rewinds and static photos? Distances ranged from 0 to 206,155 meters. For analysis, these distances were logarithmically scaled and normalized into six distance groups to mimic human categorization of distances into generalizations like ``close,'' ``far,'' or ``very far.'' When normalized: Group 0 corresponded to a range of 0--10 meters from the participant's home location; Group 1 corresponded to a range of 10--100 meters; Group 2 corresponded to a range of 100--1,000 meters; Group 3 corresponded to a range of 1,000--10,000 meters; Group 4 corresponded to a range of 10,000--100,000 meters and Group 5 corresponded to a range of 100,000--1,000,000 meters from the participant's home location.  
	\item Visit frequency: do participants react differently to Rewinds and static photos depicting locations they visit frequently compared to locations they rarely visit? We showed locations that were either ``most frequently visited,'' ``least frequently visited,'' or randomly selected.
	\item GPS usage: do participants who use GPS navigation regularly react differently to Rewinds and static photos compared to participants who don't use GPS navigation regularly?
	\item Traveler type: do attitudes toward Rewinds and static photos differ depending on whether the participant identifies as a frequent traveler?
	\item Companions: how does the presence of a traveling companion affect the participant's attitude toward Rewinds or static photos?
\end{enumerate}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Within-Subject Study}
\subsubsection{Memorability}
Within-subject comparison in our study reveals that participants are significantly better at recalling semantic memory of a particular location with Rewinds than with static street-view photos, $\chi^2(1, 356) = 96.987$, $p < 0.01$. When being presented with a street-side photo, participants were able to recognize places visited 50\% of the time; yet recognition rate of locations reached 97\% when participants played accompanying Rewinds. It should not be too surprising that cinematic media like a Rewind would provide participants with more contextual meanings and retrival cues for semantic processing, as more than one participant have expressed their contentment after wathing \textit{Rewind} videos by claiming ``I recognize the photo now!'' when their initial attempts with static images were fruitless.

It could possibly be inferred that participants are likely to benefit more from \textit{Rewind} videos than static street-view images at recalling episodic memory. Although participants were not explicitly asked about trip remembrance when presented with static images, location rememberance from static images may serve as an upper-bound estimation of how much an individual could recall about a particular trip. %- tbc
\subsubsection{Believability and Desirability}
Although not statistically significant, %TODO TODO TODO TODO TODO

\subsection{Between-Subject Study}
\begin{table*}[t]
	\centering
	\caption{My caption}
	\small
	\label{within-subject-table}
	\begin{tabular}{c|c|c|c|c|c|c|c|c|}
		\cline{2-9}
		& \multicolumn{4}{c|}{Memorability} & \multicolumn{2}{c|}{Belivability} & \multicolumn{2}{c|}{Desirability} \\ \cline{2-9} 
		& \begin{tabular}[c]{@{}c@{}}video\\  location\end{tabular} & \multicolumn{2}{c|}{\begin{tabular}[c]{@{}c@{}}photo\\ location/trip\end{tabular}} & \begin{tabular}[c]{@{}c@{}}video\\ trip\end{tabular} & photo & video & photo & video \\ \hline
		\multicolumn{1}{|c|}{N} & 177 & \multicolumn{2}{c|}{179} & 177 & 86 & 128 & 179 & 177 \\ \hline
		\multicolumn{1}{|c|}{\%} & 96.61\% & \multicolumn{2}{c|}{49.72\%} & 53.67\% & 77.91\% & 79.69\% & 10.61\% & 14.69\% \\ \hline
		\multicolumn{1}{|c|}{p-val} & \multicolumn{2}{c|}{\begin{tabular}[c]{@{}c@{}}$\chi^2(1, 356) = 96.987$\\ p = 0.0001\end{tabular}} & \multicolumn{2}{c|}{\begin{tabular}[c]{@{}c@{}}$\chi^2(1, 356) = 0.41$\\ p = 0.2611\end{tabular}} & \multicolumn{2}{c|}{\begin{tabular}[c]{@{}c@{}}$\chi^2(1, 214) = 0.02$\\ p = 0.4434\end{tabular}} & \multicolumn{2}{c|}{\begin{tabular}[c]{@{}c@{}}$\chi^2(1, 356 ) = 0.995$\\ p = 0.1593\end{tabular}} \\ \hline
	\end{tabular}
\end{table*}
% 2nd table %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

% shall we change photo/rewind video rememberance to location/trip rememberance? april - i define photo/rewind remembrance below. a photo and a rewind can have both location and trip remembrance. for instance from a photo you might remember the location but not the trip. 
\subsection{Memorability}
Rewinds, and digitally-mediated memories in general, are meant to engage a person's memory. To find out what features make a digital memory memorable, we examined participant responses to whether or not they were able to remember the location shown in a photo (semantic memory) or recall the trip portrayed by a Rewind (episodic memory). Photo memorability is determined by whether or not the participant remembered the location depicted in a photo. Because a Rewind contains many locations, it would be unfair to consider the memorability of Rewind based on whether the participant remembers any locations in the Rewind. So, Rewind memorability is determined by whether or not the participant remembered and could distinguish the particular trip portrayed in a Rewind. In this section, we will look at differences in photo and Rewind memorability based on the five metrics discussed previously.

%TODO heather - didn't do the stats for video location vs photo locations, not sure if it is necessary? - photo vs video stats?
%TODO will add this part back in within-group comparison

@@

\subsubsection{Distance}
% only affect location rememberance
% 0 - 100%, 1 - 75%, 2 - 72%, 3 - 44%, 4 - 52%, 5 - 10%, $\chi^2(5,179) = 18.106$, p < 0.01
The distance away from the participant’s home had differing effects on photo memorability and Rewind memorability (Figure~\ref{fig:distancememory}).

@@
